 

### 2023-05-22 markdown/methods.Rmd
try catch for combat implemented
next --> implement for exp_set

#### Batch correction 
Since mass spectrometry experiments are highly biased by several factors, such as sample processing, the technician, manufacturer and production of the device (chromatography column, mass-spectrometer), the reagents etc., working in multiple batches introduces further variance in the protein data. This variance would overshadow the biological differences across sample types. Before attempting to remove batch effects it is recommended to elucidate the variance with a dimensionality reduction technique which will be explained in the next chapter. To remove the technical factor also called non-biological experimental variation the R package sva \citep{Leek2012} with the ComBat function \citep{Johnson2007} was used. The expression of a gene $g$ for the batch $m_i$ and sample $j$ is defined as  $Y_{ijg} = a_{g} + X\beta_{g} + \gamma_{ig} + \delta_{ig}\epsilon_{ijg}$, whereas $\gamma$ and $\delta\epsilon$ denote for the systematic error caused by the batch. In order to remove the systematic error the bayesian framework shrinks the effect by pooling information over the features $g$. Ideally the effects $a$ and $\beta$ remain as natural as they are and should reflect the true biological difference between experimental obtained samples or groups. ComBat standardizes the data by a Z-transformation before attempting to remove the batch effect. By the method of moments the two batch effect parameters $\gamma$ and $\delta\epsilon$ were estimated and can be subtracted from each value for the particular feature.

>>>>

#### Batch correction 
Since mass spectrometry experiments are highly biased by several factors, such as sample processing, the technician, manufacturer and production of the device (chromatography column, mass-spectrometer), the reagents etc., working in multiple batches introduces further variance in the protein data. This variance would overshadow the biological differences across sample types. Before attempting to remove batch effects it is recommended to elucidate the variance with a dimensionality reduction technique which will be explained in the next chapter. To remove the technical factor also called non-biological experimental variation the R package sva \citep{Leek2012} with the ComBat function \citep{Johnson2007} was used. The expression of a gene $g$ for the batch $m_i$ and sample $j$ is defined as  $Y_{ijg} = a_{g} + X\beta_{g} + \gamma_{ig} + \delta_{ig}\epsilon_{ijg}$, whereas $\gamma$ and $\delta\epsilon$ denote for the systematic error caused by the batch. In order to remove the systematic error the bayesian framework shrinks the effect by pooling information over the features $g$. Ideally the effects $a$ and $\beta$ remain as natural as they are and should reflect the true biological difference between experimental obtained samples or groups. ComBat standardizes the data by a Z-transformation before attempting to remove the batch effect. By the method of moments (= the mean and the variance) the two batch effect parameters $\gamma$ and $\delta\epsilon$ were estimated and can be subtracted from each value for the particular feature driving statistical moments at a comparable level between batches. 
Furthermore the ComBat function allows the user to pass a desired effect (the actual biological variance) to be crystallized out from the non-biological variance. This is done by a constructing a design matrix which is also passed to the bayesian framework and can be interpreted as the apriori information about our dataset.
